---
title: "Compare ESD and ESG Quantitative Benchmarks"
author: 
output:
  pdf_document: default
  word_document: default
header-includes:
 - \usepackage[width=\textwidth]{caption}
always_allow_html: yes
csl: ../citations/citations/apa.csl
bibliography: ../citations/citations/citations.bib
link-citations: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
```

# Introduction

The development of Ecological Sites (ES) and their descriptions (ESD's) represent an enormous effort on behalf of the Natural Resources Conservation Service (NRCS) (CITE : BESTELMEYER). As mentioned in section XX, they do not yet form a continuous coast to coast, nor even across the field office data set (CITE : TWIDWELL). In order to allay the significant amounts of effort required to develop these, and to make land management decisions in the interim alternative classification systems have been proposed (CITE). One system which incorporates the NRCS ESDs, and which provides continuous coverage across our study area, are the Ecological Site Groups (CITE). 

The Ecological Site Groups developed largely by United States Geological Survey (USGS) researchers at the Southwest Biological Science Center in Moab [Utah], with assistance from BLM Colorado State Office Staff, are meant to both bridge the spatial gap in ESD development, and to provide a framework for  land management decisions at a scale larger than that which generally occurs at a BLM Field Office. The Ecological Site Groups were developed by grouping together similar ESD's, akin to the strictly interpretive approach - which seeks to reduce conceptual fineness - in section XX, and which will respond in a similar fashion to management actions (@duniway2016generalizing). To these initial groups the field data which largely informed the ESD creation, most notably soil physical parameters, and with the use of an objective and quantitative approaches, which make use of simple *Machine Learning (ML)* method *Random Forest* **section XX** were used to identify recurring themes in the dataset into groups, and then project these onto a 'map' which covers the Colorado Plateau. 

While for most land management decisions at the UFO ESD's, when available, represent the best available scientific evidence upon which to inform decisions,  ESG's provide the best alternative for much of the field office, and in the future are likely to maintain influence for large scale decisions at the UFO. Herein we address questions regarding the similarity of estimates arising from ESG's and ESD's. In this report, we are using ESD's for the *reference state benchmarks* which they contain, in other words quantitative goals which we seek to compare land to, and seek to visualize the relationship between the standards of ESD's and ESG's. We also hope to understand which ecological sites are contained in which ES groups, in other words, where do our groups map out to?. 


> "QUOTE WILL GO HERE SUMMARIZING WHAT ESGS ARE ALL ABOOT"
>  
> `r tufte::quote_footer('--- Nauman et al. 2022')`

# Methods

Here we determine how similar they are to the ESD's for which we have verified AIM plots in. In order to determine if ESG's may serve as a drop-in-substitute for areas of the field office which do not have completed ESD's...


```{r}
library(tidyverse)
library(sf)
library(terra)
library(bipartite)

set.seed(51)
```

```{r Import Quantitative Fractional Cover Benchmarks}
praw <- '../data/raw'
f <- list.files(praw, recursive = T)

ESD_quant <- read.csv(file.path(praw, f[grep('ESD_Quantitative',f)] )) %>% 
  rowwise() %>% 
  mutate(MEAN = sum(LOWER,  UPPER, na.rm = T)/2) 

ESG_cover <- read.csv(file.path(praw, f[grep('FractionalCover', f)] )) %>% 
  rowwise() %>% 
  mutate(GRASS = sum(Sacaton, C3_Perennial_Grass, C4_Perennial_Grass, Grass),
         SHRUB = sum(Saltbush, Shadescale, Shrub)) %>% 
  select(Climate:SoilGeomorphicUnits, GRASS, SHRUB, FORB = Forb, TREE = Tree) %>% 
  pivot_longer(cols = GRASS:TREE, names_to = 'COVER_TYPE', values_to = 'MEAN') 

```

```{r Import AIM plot locations subset to ESD verified and find their ESG via spatial operations}

ecosites <- read.csv(file.path(praw, f[grep('Plot_Tracking',f)] )) %>% 
  filter(ECO.SITE.MATCHED == T)

esg <- rast(file.path(praw, f[grep('ESGs.*tif$',f)]))

ecosites <- st_read('/media/sagesteppe/ExternalHD/UFO_cartography/AIM_plots_outcomes/plots_outcomes.shp', 
                 quiet = T) %>% # let us grab the spatial data from this real quick
  select(PLOT.ID = Plot_ID) %>% 
  st_transform(26912) %>% 
  st_buffer(27.5) %>% 
  right_join(., ecosites, by = 'PLOT.ID')  %>% 
  vect() %>% 
  project(., crs(esg))

mode <- function(codes){as.numeric(which.max(tabulate(codes)))}

esg2plots <- extract(esg, ecosites, method = 'simple', weights = T, ID = T) %>% 
  group_by(ID) %>% 
  summarise(Value = mode(ESG)) %>% 
  cbind(PLOT.ID = ecosites$PLOT.ID, .)

esg2plots <- data.frame(levels(esg)) %>% # extract lookup table from raster
  right_join(., esg2plots, by = 'Value') %>% 
  select(ESG, PLOT.ID)
ecosites <- as.data.frame(ecosites) %>%
  select(ECO.SITE, PLOT.ID)

ESD2ESG_map <- left_join(esg2plots, ecosites, by = 'PLOT.ID')
  
rm(mode, esg, f, praw, esg2plots, ecosites)
```


```{r}

ESD2ESG_wide <- ESD2ESG_map %>% 
  mutate(
    ESG = str_replace_all(ESG, ' ', "_"),
    ESG = str_replace(ESG, ',_', '_and_'),
    ESG = str_replace(ESG, 'Semiarid_Warm', 'SAW'),
    ESG = str_replace(ESG, 'Semiarid_Cool', 'SAC'), 
    ESG = str_replace(ESG, 'Arid_Warm_', 'AW')) %>% 
  select(-PLOT.ID) 

ob <- with(ESD2ESG_wide, table(ESG, ECO.SITE))
EsdEsg_network <- apply(as.matrix.noquote(ob),2,as.numeric)
rownames(EsdEsg_network) <- rownames(ob)

mod <- computeModules(EsdEsg_network)
plotModuleWeb(mod)


NLINteractions <- networklevel(EsdEsg_network, index = c('linkage density', 
                                       'weighted nestedness', 
                                       'H2'))

# Higher Level refers to the ESD's , Lower level to the ESGs
GLInteractions <- grouplevel(EsdEsg_network,
             index = c(
               'number of species', # no expectations is what went in
               'mean number of links', # for ESD expect 1, ESG > 1
               'mean number of shared partners', # for ESG > 1, for ESD still greater than 1 as many can pool into one ESG
               'niche overlap', # should be very low for ESG, should be higher for ESD, but still relatively low. 
               'C score' # should be VERY HIGH for upper, low for lower.  
               ))

plotweb(mod@moduleWeb)

SLInteractions <- specieslevel(EsdEsg_network, index = c('degree',  # number of links per species, ESD should be 1, ESG > 1
  'species strength',
  'effective partners',
  'nestedrank'), level = 'both')

modules <- module2constraints(mod)
module_membership <- data.frame(c(rownames(EsdEsg_network), colnames(EsdEsg_network)), modules) 
colnames(module_membership) <- c("vertices", "modules") 

module_membership <- module_membership %>% 
  mutate(vertices = str_remove_all(vertices, ','),  
         vertices = str_replace(vertices, '_Finer_Uplands_and_Clay_Uplands', '_Finer_and_Clay_Uplands'),
         vertices = str_replace(vertices, 'Sandy_Uplands_and_Loamy_Uplands', 'Sandy_and_Loamy_Uplands')
         )

rm(ob, mod, EsdEsg_network, modules, ESD2ESG_map)
```

```{r Facet wrap density plots of ESD values with geom point ESG value}

ESD_quant <- ESD_quant %>% 
  select('CONCEPT' = ECO.SITE, COVER_TYPE, MEAN) %>% 
  mutate(MODEL = 'ESD')

ob <- ESG_cover %>% 
  unite('ESG', Climate:SoilGeomorphicUnits, sep = '_-_') %>% 
  mutate(
    ESG = str_replace_all(ESG, ' ', "_"),
    ESG = str_remove(ESG, 'No_Climate_Group_-_'),
    ESG = str_replace(ESG, 'Semiarid_warm', 'SAW'),
    ESG = str_replace(ESG, 'Semiarid_Cool', 'SAC'), # do need to go back to clean up names
    ESG = str_replace(ESG, 'Arid_Warm_', 'AW'),  # Finer_and_Clay_Uplands
    ESG = str_replace(ESG, 'Arid_Warm_', 'Finer_and_Clay_Uplands'), 

    
    
    
    ) %>% # of some of these strata
  mutate(MODEL = 'ESG') %>% 
  rename(CONCEPT = ESG) %>% 
  bind_rows(ESD_quant) %>% 
  right_join(., module_membership, by = c('CONCEPT' = 'vertices')) %>% 
  mutate(MEAN_X = MEAN, MEAN_Y = MEAN) %>% 
  filter(COVER_TYPE %in% c('FORB', 'GRASS', 'SHRUB', 'TREE'))

ggplot(ob, aes(MEAN_X, MEAN_Y, color = MODEL)) +
  geom_jitter() +
  facet_grid(rows = vars(modules), cols = vars(COVER_TYPE)) +
  theme_bw()


```



# Results

We run multiple tests of correlation coefficients to determine similarity between these two data sets. ...

# Conclusions

The relationship between the novel ESG's and the ESD's...


