---
title: "Compare ESD and ESG Quantitative Benchmarks"
author: 
output:
  pdf_document: default
  word_document: default
header-includes:
 - \usepackage[width=\textwidth]{caption}
 - \usepackage{float}
 - \floatplacement{figure}{H}
always_allow_html: yes
csl: ../citations/citations/apa.csl
bibliography: ../citations/citations/citations.bib
link-citations: yes
---

```{r global options, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(fig.pos = 'H')
```

# Introduction

The development of Ecological Sites (ES) and their descriptions (ESD's) represent an enormous effort on behalf of the Natural Resources Conservation Service (NRCS) (@bestelmeyer2015national). As mentioned in section XX, they do not yet form a continuous coast to coast, nor even across the field office data set (@twidwell2013national). In order to allay the significant amounts of effort required to develop these, and to make land management decisions in the interim alternative classification systems have been proposed (@nauman2022quantitative). One system which incorporates the NRCS ESDs, and which provides continuous coverage across our study area, are the Ecological Site Groups (@nauman2022quantitative). 

The Ecological Site Groups developed largely by United States Geological Survey (USGS) researchers at the Southwest Biological Science Center in Moab Utah, with assistance from BLM Colorado State Office Staff, are meant to both bridge the spatial gap in ESD development, and to provide a framework for  land management decisions at a larger scale than that which generally occurs at a BLM Field Office. The Ecological Site Groups were developed by grouping together similar ESD's, akin to the strictly interpretive approach - which seeks to reduce conceptual fineness - in section XX, and which will respond in a similar fashion to management actions (@duniway2016generalizing). To these initial groups the field data which largely informed the ESD creation, most notably soil physical parameters, and with the use of an objective and quantitative approaches, which make use of simple *Machine Learning (ML)* method *Random Forest* **section XX** were used to identify recurring themes in the dataset into groups, and then project these onto a 'map' which covers the Upper Colorado River Basin. 

While for most land management decisions at the UFO ESD's, when available, represent the best available scientific evidence upon which to inform decisions,  ESG's provide the best alternative for much of the field office, and in the future are likely to be influential for large scale decisions at the UFO. Herein we address questions regarding the similarity of estimates arising from ESG's and ESD's. In this report, we are using ESD's for the *reference state benchmarks* which they contain, in other words quantitative goals which we seek to compare land to, and seek to visualize the relationship between the standards of ESD's and ESG's. 

The goals of this section are to 1) determine which ESD's constitute the ESG's across the UFO? And 2) visualize the similarity in benchmarks within an ESG and the ESD's which it contains. A short investigation of how different methods of extracting the spatial dataset to the ground truthed data, as well as a comparision of the accuracy of the predicted ESG's limited to the UFO area and ground truthed plots, is also included. 

In the publication of Nauman et al. 2022 the accuracy of the final product was calculated as being 57.4% (@nauman2022quantitative). However, In our expereince certain groups are often more difficult to reliably predict, and without sufficient testing sample sizes many spatial products will score higher, or lower, than these averages in areas with/without these features. 

# Methods

```{r}
url <- urlshorteneR::isgd_LinksShorten(longUrl = "https://www.sciencebase.gov/catalog/item/5f76119e82cef8d1839c681d", showRequestURL = F)
```

The mean fractional cover of vegetation was manually transcribed from Table 3 of Appendix A from Nauman et al. 2022. The spatial data product, a gridded surface of predicted ESG's, which was the outcome of the study was accessed from sciencebase.gov catalog on (Dec. 16, 2022, `r url`), three additional layers predicting soil geomorphological groups and another predicting climate zone were also downloaded. ESD quantitative benchmarks values, and AIM plot locations were cleaned in previous work (Section XX ESD Completion). All analyses were performed in R version 4.2.2 using RStudio on Linux Ubuntu 22.04 LTS (@r2022language, @sobell2015practical, @rstudio). 

```{r}
library(tidyverse)
library(sf)
library(terra)
library(bipartite)

source('../../UFO_elements_of_style/scripts/ggplot_themes.R')
set.seed(51)

praw <- '../data/raw'
f <- list.files(praw, recursive = T)

rm(url)
```


```{r Import AIM plot locations subset to ESD verified and find their ESG via spatial operations}

ecosites <- read.csv(file.path(praw, f[grep('Plot_Tracking',f)] )) %>% 
  filter(ECO.SITE.MATCHED == T)

esg <- rast(file.path(praw, f[grep('ESGs.*tif$',f)]))

ecosites <- st_read('/media/sagesteppe/ExternalHD/UFO_cartography/AIM_plots_outcomes/plots_outcomes.shp', 
                 quiet = T) %>% # let us grab the spatial data from this real quick
  select(PLOT.ID = Plot_ID) %>% 
  st_transform(26912) %>% 
  st_buffer(27.5) %>% 
  right_join(., ecosites, by = 'PLOT.ID')  %>% 
  vect() %>% 
  project(., crs(esg))

mode <- function(codes){as.numeric(which.max(tabulate(codes)))}

esg2plots <- extract(esg, ecosites, method = 'simple', weights = T, ID = T) %>% 
  group_by(ID) %>% 
  summarise(Value = mode(ESG)) %>% 
  cbind(PLOT.ID = ecosites$PLOT.ID, .)

esg2plots <- data.frame(levels(esg)) %>% # extract lookup table from raster
  right_join(., esg2plots, by = 'Value') %>% 
  select(ESG, PLOT.ID) 
ecosites_mapped <- as.data.frame(ecosites) %>% 
  select(ECO.SITE, PLOT.ID) 

ESD2ESG_map <- left_join(esg2plots, ecosites_mapped, by = 'PLOT.ID') %>% 
  mutate(
    ESG = str_replace_all(ESG, ' ', "_"), 
    ESG = str_replace(ESG, ',_', '_and_'), 
    ESG = str_replace(ESG, 'Semiarid_Warm', 'SAW'), 
    ESG = str_replace(ESG, 'Semiarid_Cool', 'SAC'),  
    ESG = str_replace(ESG, 'Arid_Warm_', 'AW'))  
  
rm(esg2plots, ecosites_mapped)
```

```{r Use ideal lookup table from publication}

ref_lookup <- read.csv(file.path(praw, f[grep('Appendix',f)] )) %>% 
  select(Ecological.Site.ID, Final.Soil.Geomorphic.Unit, Climate.Zone) %>% 
  filter(Ecological.Site.ID %in% ESD2ESG_map$ECO.SITE) %>% 
  mutate(
    Climate.Zone = str_replace(Climate.Zone, 'Semiarid_Warm', 'SAW'), 
    Climate.Zone = str_replace(Climate.Zone, 'Semiarid_Cool', 'SAC'),  
    Climate.Zone = str_replace(Climate.Zone, 'Arid_Warm', 'AW'),  
    Final.Soil.Geomorphic.Unit = snakecase::to_parsed_case(Final.Soil.Geomorphic.Unit) 
    ) %>% 
  mutate(Final.Soil.Geomorphic.Unit = case_when( 
    Climate.Zone == 'SAW' & Final.Soil.Geomorphic.Unit %in% c('Shallow', 'Deep_Rocky') ~ 'Shallow_and_Deep_Rocky', 
    TRUE ~ Final.Soil.Geomorphic.Unit
  )) %>% 
  unite('ESG', Climate.Zone:Final.Soil.Geomorphic.Unit, sep = '_-_') 

ESG_present <- nrow(ref_lookup) 

rm(f, praw)
```


### Test Raster Extraction Methods 


```{r test raster extraction methods}

ecosites <- st_read('/media/sagesteppe/ExternalHD/UFO_cartography/AIM_plots_outcomes/plots_outcomes.shp', 
                 quiet = T) %>% # let us grab the spatial data from this real quick
  select(PLOT.ID = Plot_ID) %>% 
  st_transform(26912) 

ESG4simulation <- filter(ESD2ESG_map, ECO.SITE %in% ref_lookup$Ecological.Site.ID) %>% 
  select(-ESG) %>% 
  left_join(., ecosites, by = 'PLOT.ID') %>% 
  st_as_sf() %>% 
  vect() %>% 
  project(., crs(esg))

ESG4simulation_buff <- ESG4simulation %>% 
  st_as_sf() %>% 
  st_transform(26912) %>% 
  st_buffer(27.5) %>% 
  vect() %>% 
  project(., crs(esg))

known <- st_as_sf(ESG4simulation) %>% 
  st_drop_geometry()

## Extract Values for the plot center ##
bN <- extract(esg, ESG4simulation, method = 'simple') %>% 
  mutate(test = 'point') %>% 
  cbind(known)

## Extract Values for the buffer plots
bY <- extract(esg, ESG4simulation_buff, method = 'simple',  bind = T) 
bY <- extract(esg, ESG4simulation_buff, method = 'simple') %>% 
  group_by(ID) %>% 
  summarise(Value = mode(ESG)) %>% 
  cbind(known)
bY <- data.frame(levels(esg)) %>% 
  right_join(., bY, by = 'Value') %>% 
  select(-Value) %>% 
  mutate(test = 'polygon') 

test_results <- rbind(bN, bY) %>% 
  mutate(
    ESG = str_replace_all(ESG, ' ', "_"),
    ESG = str_replace_all(ESG, ',', "_and"),
    ESG = str_remove(ESG, 'No_Climate_Group_-_'),
    ESG = str_replace(ESG, 'Semiarid_Warm', 'SAW'),
    ESG = str_replace(ESG, 'Semiarid_Cool', 'SAC'), 
    ESG = str_replace(ESG, 'Arid_Warm', 'AW'),
    ) %>% 
  select(-ID, -PLOT.ID) %>% 
  mutate(RESULT = if_else(ESG %in% ref_lookup$ESG & ECO.SITE %in% ref_lookup$Ecological.Site.ID, 1, 0)) 

p2p <- test_results %>% 
  group_by(test) %>% 
  summarize(Prop = sum(RESULT)/n()) # polygon gives slightly better results

rm(bY, bN, ESG4simulation, ESG4simulation_buff, ecosites)
```


To test whether we were extracting and processing the correct ESG from the gridded surface two comparisons of raster extraction methods were attempted. The first (hereafter: 'polygon'), which is typically performed at UFO, is to buffer the point to represent the actual area of the entire AIM plot and extract to that area, and choose the categorical class with the most cells per value (in other words the statistical mode). The second (herafter: 'point') option is to extract values from the gridded surface directly to the point geometry. The idea behind this is that the pixels from the gridded surface featured a 'splotchy' characteristic, a byproduct of not performing spatial operations to clean up the predictions (e.g. 'Focal Statistics'); oftentimes these central isolated cells are not artifacts of analysis, but rather true points which are swamped by the adjacent cells. The slightly higher performing method, 'polygon', was used for the duration of the analyses. 

### Test Local Accuracy of Raster Dataset 

To determine whether the raster dataset gives a higher or lower performance accuracy in the UFO portion of it's range the `r nrow(known)` AIM plots in  `r nrow(ref_lookup)` ESD's which are known to map directly to `r length(unique(ref_lookup$ESG))` ESG's were utilized. These plots were extracted from the ESG gridded surface and the proportion of each ESG which was correctly and incorrectly mapped were calculated.  

```{r Proportion of Correct ESG matching}

# The percent of ESG's which were classified as such, and the percent of those mappings which
# are accurate. 
esg_correct <- test_results %>% 
  filter(test == 'polygon') %>% 
  group_by(ESG) %>% 
  mutate(Correct = sum(RESULT), 
         Total = n(),
         Incorrect = Total - Correct) %>% 
  distinct(ESG, Correct, Incorrect, Total, .keep_all = F) %>% 
  pivot_longer(cols = c(Correct, Incorrect), values_to = 'Response_V', names_to = 'Response')

number_incorrect <- esg_correct %>% 
  filter(Response == 'Incorrect', Total == Response_V) %>% 
  nrow()
number_correct_esg <- esg_correct %>% 
  filter(Response == 'Correct', Total == Response_V) %>% 
  nrow()

# Proportion of which plots per ESD are correctly identified by the raster surface. 
esd_correct <- test_results %>% 
  filter(test == 'polygon') %>% 
  group_by(ECO.SITE) %>% 
  mutate(Correct = sum(RESULT), 
         Total = n(),
         Incorrect = Total - Correct) %>% 
  distinct(ECO.SITE, Correct, Incorrect, Total, .keep_all = F) %>% 
  pivot_longer(cols = c(Correct, Incorrect), values_to = 'Response_V', names_to = 'Response')

esd_corr_plots <- ggplot(esd_correct, aes(y = Response_V, x = ECO.SITE, fill= Response)) +
    geom_bar(position="stack", stat="identity") +
    theme(strip.background = element_blank(), 
          panel.background = element_blank(),
          plot.title = element_text(hjust = 0.5)
          ) +
    scale_fill_manual('Identification:', 
                      values = c('Correct' = '#88A47C', 'Incorrect' = '#E6E2C3')) +
  scale_x_discrete(guide = guide_axis(n.dodge=4)) +
  labs(title = 'Proportion of Plots mapped to their ESG', y = 'Number of Plots', x = 'Ecological Site') 

# ESD's differ in the proportion of plots which may be mapped over to ESG. 
# In general our results here echo the paper, 1/3 of records will be incorrectly mapped from 
# the single output prediction. 

prop_D_map2_G <- esd_correct %>% 
  filter(Response == 'Correct') %>% 
  ungroup() %>% 
  summarize(prop_correct = sum(Response_V) / sum(Total))
            
rm(theme_boxplot, theme_prop_bar, esg, esg_correct, test_results, esd_correct)
```


### Determine ESD and ESG Groupings 

To inform a strategy for mapping individual ESDs directly to ESG's a bipartite network was created using all AIM plots with verified ESD's,  all statistics (**not reported in text**) were calculated using the package 'bipartite' (@dormann2008bipartite). This approach was deemed neccessary given how many erroneous relationships were returned by the ESG extraction method. 

```{r Create Networks to show relationships between pixels and classification.}

ESD2ESG_wide <- ESD2ESG_map %>% 
  select(-PLOT.ID) 

ob <- with(ESD2ESG_wide, table(ESG, ECO.SITE))
EsdEsg_network <- apply(as.matrix.noquote(ob),2,as.numeric)
rownames(EsdEsg_network) <- rownames(ob)

orig_mod <- computeModules(EsdEsg_network)

# Higher Level refers to the ESD's , Lower level to the ESGs

NLINteractions <- networklevel(EsdEsg_network, index = c(
  'linkage density', 
  'weighted nestedness', 
  'H2'))

GLInteractions <- grouplevel(EsdEsg_network, index = c(
  'number of species', # no expectations is what went in
  'mean number of links', # for ESD expect 1, ESG > 1
  'mean number of shared partners', # for ESG > 1, for ESD still greater than 1 as many can pool into one ESG
  'niche overlap', # should be very low for ESG, should be higher for ESD, but still relatively low. 
  'C score' # should be VERY HIGH for upper, low for lower.  
  ))

SLInteractions <- specieslevel(EsdEsg_network, index = c(
  'degree',  # number of links per species, ESD should be 1, ESG > 1
  'species strength', 
  'effective partners',
  'nestedrank'))

rm(ob, EsdEsg_network, NLINteractions, GLInteractions, SLInteractions)
```

All observed ESD's were matched to ESG's in an incremental fashion. 1) ESD's which were used directly in the creation of ESG's were removed based upon a noted association in Appendix 6 of Naumen et al. 2022. 2) ESD's for which only a single AIM plot existed, had the ESG which values were extracted from it listed as it's ultimate mapped association. 3) ESD's with greater than 1 AIM plots associated with them, and which had all plots match an ESG were classified as such. 4) ESD's with over 65%, our local accuracy of the gridded surface, of their AIM plots mapping to a single ESG. 

```{r Create a clean network of ESDs which map to an ESG}

# ' the clean network '

cleanNET1 <- ESD2ESG_wide %>% # these in ESG table
  ungroup() %>% 
  filter(ECO.SITE %in% ref_lookup$Ecological.Site.ID) %>% 
  select(ECO.SITE) %>% 
  left_join(., ref_lookup, by = c('ECO.SITE' = 'Ecological.Site.ID'))

cleanNET2 <- ESD2ESG_wide %>% # these ESDs only a single plot, take extracted ESG
  ungroup() %>% 
  filter(!ECO.SITE %in% ref_lookup$Ecological.Site.ID) %>% 
  group_by(ECO.SITE) %>% 
  filter(n()== 1)

cleanNET3 <-  ESD2ESG_wide %>% #these ESG's match perfectly between all plots 
  ungroup() %>% 
  filter(!ECO.SITE %in% ref_lookup$Ecological.Site.ID) %>% 
  group_by(ECO.SITE) %>%
  filter(n() >= 2) %>% 
  
  add_tally(name = 'NO_ESD') %>% #these ESG's match perfectly between all plots 
  group_by(ECO.SITE, ESG) %>% # within the ESD ! F048AY918CO 4/4, R034XY403CO 5/5 
  add_tally(name = 'ESD_in_ESG') %>% 
  filter(NO_ESD == ESD_in_ESG) %>% 
  select(-NO_ESD, -ESD_in_ESG)

cleanNET4 <- ESD2ESG_wide %>%  # over 65% of plots map to same esg
  ungroup() %>% 
  filter(!ECO.SITE %in% ref_lookup$Ecological.Site.ID) %>% 
  group_by(ECO.SITE) %>%
  filter(n() >= 2) %>% 
  
  add_tally(name = 'NO_ESD') %>% #these ESG's match perfectly between all plots 
  group_by(ECO.SITE, ESG) %>% # within the ESD ! F048AY918CO 4/4, R034XY403CO 5/5 
  add_tally(name = 'ESD_in_ESG') %>% 
  filter(NO_ESD != ESD_in_ESG) %>% 
  
  filter(ESD_in_ESG / NO_ESD > 0.65) %>% 
  distinct() %>% 
  slice(rep(1:n(), each = NO_ESD)) %>%  # now have to remove these from the next set
  select(-NO_ESD, -ESD_in_ESG)

cleanNet <- rbind(cleanNET1, cleanNET2, cleanNET3, cleanNET4)

ob <- with(cleanNet, table(ESG, ECO.SITE))
EsdEsg_network <- apply(as.matrix.noquote(ob),2,as.numeric)
rownames(EsdEsg_network) <- rownames(ob)
clean_mod <- computeModules(EsdEsg_network)
```


```{r Create subset Network for remaining ESDs which needan ESG}

# ' the dirty network'

dirtyNet <- ESD2ESG_wide %>% 
  ungroup() %>% 
  filter(!ECO.SITE %in% ref_lookup$Ecological.Site.ID) %>% # the groups of these ESD's are known #/ 157 plots
  
  group_by(ECO.SITE) %>% # these ESD's had only a single plot, and we will just #/ 8 plots
  filter(n() >= 2) %>% # accept whichever ESG it is mapped to. 
  
  add_tally(name = 'NO_ESD') %>% #these ESG's match perfectly between all plots 
  group_by(ECO.SITE, ESG) %>% # within the ESD ! F048AY918CO 4/4, R034XY403CO 5/5 
  add_tally(name = 'ESD_in_ESG') %>% 
  filter(NO_ESD != ESD_in_ESG) %>% 
  
  filter(ESD_in_ESG / NO_ESD < 0.65) %>%  # sites with > 65% of plots to an ESG will be assigned as such. 
  filter(!ECO.SITE %in% c('R048AY238CO', 'R048AY306UT')) # R048AY238CO, R048AY306UT


ob <- with(dirtyNet, table(ESG, ECO.SITE))
EsdEsg_network <- apply(as.matrix.noquote(ob),2,as.numeric)
rownames(EsdEsg_network) <- rownames(ob)
dirt_mod <- computeModules(EsdEsg_network)

rm(ob, EsdEsg_network)
```

Following the removal of these ESDs from the initial dataset, the three Soil Geomorphologic Units (SGU) surfaces, and the Climate Zone surface were used to match the remaining ESDs to an ESG. Each of the remaining points had values extracted from each layer of the SGU surfaces, each representing the classification prediction from the three top performing classifier models. The most commonly occurring ESG (in others words the statistical mode), was then calculated considering each of the SGU layers per point, across all points in the ESD simultaneously. All climate zones were also extracted to the points, and the most commonly occurring climate zone per ESD was selected as the climate zone to classify these sites in the SGU framework, thus forming the ESG. 

```{r Classify Remaining ESDs to an ESG based on the most commonly observed one across all Plots }
praw <- '../data/raw'
f <- list.files(praw, recursive = T)

sgu1  <- rast( file.path(praw, f[grep('SGU_1st.*tif$',f)]) )
sgu2  <- rast( file.path(praw, f[grep('SGU_2nd.*tif$',f)]) )
sgu3  <- rast( file.path(praw, f[grep('SGU_3rd.*tif$',f)]) )
c_zones <- rast( file.path(praw, f[grep('Zones.*tif$', f)]) )

ecosites <- st_read('/media/sagesteppe/ExternalHD/UFO_cartography/AIM_plots_outcomes/plots_outcomes.shp', 
                 quiet = T) %>% # let us grab the spatial data from this real quick
  select(PLOT.ID = Plot_ID) %>% 
  st_transform(26912) %>% 
  st_buffer(27.5) %>% 
  right_join(ESD2ESG_map, ecosites, by = 'PLOT.ID') %>% 
  filter(ECO.SITE %in% dirtyNet$ECO.SITE) %>% 
  vect() %>% 
  project(., crs(sgu1)) 

cz_lkp <- data.frame(levels(c_zones)) 
clim_zones <- extract(c_zones, ecosites, method = 'simple') %>% 
  group_by(ID) %>% summarise(ClimZone = mode(ClimZone)) %>% # mode of original input dataset
  left_join(., cz_lkp, by = c('ClimZone' = 'Value')) %>% 
  select(ID, ClimZone = ClimZone.y)  %>% 
  cbind(., PLOT.ID = st_as_sf(ecosites)) %>% 
  select(ID, ClimZone, PLOT.ID = PLOT.ID.PLOT.ID, ECO.SITE = PLOT.ID.ECO.SITE) %>% 
  ungroup() %>% 
  mutate(ClimZoneM = as.numeric(factor( # this all convoluted was much better before but then I myselfed. 
      ClimZone, levels = c("Arid Warm", "Semiarid Warm", "Semiarid Cool"), ordered = TRUE))) %>% 
  split(., .$ECO.SITE) %>% 
  map(., .f = list(. %>% mutate(ClimZoneM = mode(ClimZoneM)))) %>% 
  bind_rows() %>% 
  mutate(
    ClimZoneM = str_replace(ClimZoneM, '1', 'AW'), 
    ClimZoneM = str_replace(ClimZoneM, '2', 'SAW')
    ) %>% 
  distinct(ECO.SITE, ClimZoneM)

sgu1_ex <- extract(sgu1, ecosites, method = 'simple') %>% 
  group_by(ID) %>% summarise(SGU1 = mode(SGU)) 
sgu2 <- extract(sgu2, ecosites, method = 'simple') %>% 
  group_by(ID) %>% summarise(SGU2 = mode(SGU)) 
sgu3 <- extract(sgu3, ecosites, method = 'simple') %>% 
  group_by(ID) %>%  summarise(SGU3 = mode(SGU)) 
sgu <- data.frame(sgu1_ex, 'SGU2' =  sgu2$SGU2, 'SGU3' = sgu3$SGU3, 
                  st_as_sf(ecosites) %>% select(PLOT.ID, ECO.SITE) )

sgu <- sgu %>% 
  st_drop_geometry() %>% 
  split(., .$ECO.SITE) %>% 
  map(., .f = list(. %>% dplyr::select(2:4) %>% 
                     as.matrix() %>% 
        table(.) %>% which.max(.) %>% names(.) )) %>% 
  bind_rows() %>%  t() %>%  data.frame() %>% 
  rownames_to_column('ECO.SITE') %>% 
  rename(ESG_Code = '.')

sguvals <- data.frame(levels(sgu1)) %>% 
  mutate(Value = as.character(Value))

sgu <- left_join(sgu, sguvals, by = c('ESG_Code' = 'Value')) %>% 
  left_join(., clim_zones, by = 'ECO.SITE') %>% 
  unite('ESG', ClimZoneM, SGU, sep = '_-_') %>% 
  mutate(ESG = str_replace_all(ESG, ' ', '_'),
         ESG = str_replace(ESG, 'SAW_-_Deep_Rocky', 'SAW_-_Shallow_and_Deep_Rocky'),
         ESG = str_replace(ESG, 'SAW_-_Loamy_Uplands', 'SAW_-_Sandy_Uplands_and_Loamy_Uplands')) %>% 
  select(-ESG_Code) 

ecosites2step2 <- ecosites
rm(sgu1_ex, sgu1, sgu2, sgu3, ecosites, cz_lkp, sguvals, c_zones, dirtyNet,
   mode, clim_zones, ESD2ESG_map)
```

The results of these analyses led to the development of a draft ESD-ESG lookup table for the Uncompahgre Field Office for this project.

```{r ESD ESG Lookup table}

lookuptable <- bind_rows(distinct(cleanNet), sgu)

finalNET <- ESD2ESG_wide %>% 
  select(ECO.SITE) %>% 
  left_join(lookuptable, by = 'ECO.SITE') %>% 
  mutate(ID = 1:nrow(.))
write.csv(finalNET, '../data/processed/ESD-ESG-Lookup-Table.csv', row.names = F)

ob <- with(finalNET, table(ESG, ECO.SITE))
EsdEsg_network <- apply(as.matrix.noquote(ob),2,as.numeric)
rownames(EsdEsg_network) <- rownames(ob)
final_mod <- computeModules(EsdEsg_network)

rm(ob, EsdEsg_network, finalNET, cleanNet, ESD2ESG_wide)
```


```{r Import Quantitative Fractional Cover Benchmarks Compare Quantitative benchmarks, eval = F}

ESD_cover <- read.csv(file.path(praw, f[grep('ESD_Quantitative',f)] )) %>% 
  rowwise() %>% 
  mutate(MEAN = sum(LOWER,  UPPER, na.rm = T)/2)  %>% 
  filter(!COVER_TYPE %in% c('BAREGROUND', 'LITTER')) %>% 
  left_join(., lookuptable, by = 'ECO.SITE') %>% 
  drop_na(ESG) %>% 
  select(ECO.SITE, COVER_TYPE, MEAN) %>% 
  mutate( MODEL = 'ESD')

ESG_cover <- read.csv(file.path(praw, f[grep('FractionalCover', f)] )) %>% 
  rowwise() %>% 
  mutate(GRASS = sum(Sacaton, C3_Perennial_Grass, C4_Perennial_Grass, Grass),
         SHRUB = sum(Saltbush, Shadescale, Shrub)) %>% 
  select(Climate:SoilGeomorphicUnits, GRASS, SHRUB, FORB = Forb, TREE = Tree) %>% 
  pivot_longer(cols = GRASS:TREE, names_to = 'COVER_TYPE', values_to = 'MEAN') %>% 
  unite('ESG', Climate:SoilGeomorphicUnits, sep = '_-_') %>% 
  mutate(
    ESG = str_replace_all(ESG, ' ', "_"),
    ESG = str_remove(ESG, 'No_Climate_Group_-_'),
    ESG = str_replace(ESG, 'Semiarid_warm', 'SAW'),
    ESG = str_replace(ESG, 'Semiarid_Cool', 'SAC'), 
    ESG = str_replace(ESG, 'Arid_Warm_', 'AW'), 
    ) %>% 
#  left_join(., lookuptable) %>% 
  drop_na() %>% 
  rename(CONCEPT = ESG) %>% 
  mutate(MODEL = 'ESG')

# v <- inner_join(ESD_cover, ESG_cover, by = 'ECO.SITE')
```

```{r Facet wrap density plots of ESD values with geom point ESG value, eval = F}

ggplot(ob, aes(MEAN_X, MEAN_Y, color = MODEL)) +
  geom_jitter() +
  facet_grid(rows = vars(modules), cols = vars(COVER_TYPE)) +
  theme_bw()

```

```{r Remove variables used in methods}
rm(f, praw, lookuptable)
```

# Results

`r nrow(known)` AIM points were in the `r nrow(ref_lookup)` ESD's which were directly involved in the creation of `r length(unique(ref_lookup$ESG))` ESG's and could be used for assessing the results from the alternative raster data extraction methods and for assessing the accuracy of the ESG gridded surface in the area of analysis.  The method of extracting the ESG from the gridded surface to plot had little influence on the accuracy of the ESG. Regardless the method of buffering the point to the real size of the AIM plot was used as it slightly outperformed plot centers on the `r nrow(known)` known sites, `r round(p2p[(p2p$test == 'polygon'),]$Prop, 3)` to `r round(p2p[(p2p$test == 'point'),]$Prop, 3)`. `r nrow(known)` AIM points were in the `r nrow(ref_lookup)` ESD's which were directly involved in the creation of `r length(unique(ref_lookup$ESG))`. The ESG gridded surface was able to correctly place `r round(as.numeric(prop_D_map2_G[1,1]), 3)` plots to the appropriate ESG, rather than the 57.4% reported for the entire study area (*Figure 1*). A caveat associated with this result is that our value is likely to be biased towards more common ESG classes. We accept this bias as the scale at which we make management decisions is so great. 

```{r Proportion of plots per esd mapped correctly from the gridded data set, fig.cap="Proportion of Plots in ESDs which were included in the ESGS and should map over unequivocally to an ESG", out.width="50%"}
esd_corr_plots
rm(esd_corr_plots)
```


```{r Plot network of original Input data, fig.cap="Initial Relationship Between Field Verified ESD and ESG extracted from the gridded surface"}
plotweb(orig_mod@moduleWeb, labsize = 0.5, col.high = '#1C315E', col.low = '#227C70', col.interaction = '#E6E2C3')
```


The initial extraction of ESG to AIM points was problematic (Figure 2). 1) `r length(unique(ref_lookup$Ecological.Site.ID))` ESD's, associated with `r nrow(known)` plots, which were used directly in the creation of `r length(unique(ref_lookup$ESG))` ESG's were removed based upon a noted association in Appendix 6 of Naumen et al. 2022. 2) `r nrow(cleanNET2)` ESD's for which only a single AIM plot existed, had the `r length(unique(cleanNET2$ESG))` ESG which values were extracted from it listed as it's ultimate mapped association. These ESDs, `r unique(cleanNET2$ECO.SITE)`, mapped to the ESG's: `r unique(cleanNET2$ESG)`, refer to Figure 4 for exact matching. 3) `r length(unique(cleanNET3$ECO.SITE))` ESD's with greater than 1 AIM plots associated with them, and which had all plots match one of `r length(unique(cleanNET3$ESG))`. These ESD's, `r unique(cleanNET3$ECO.SITE)`,  were associated with `r nrow(cleanNET3)` AIM plots, and mapped to `r unique(cleanNET3$ESG)`.  4) Finally, `r length(unique(cleanNET4$ECO.SITE))` ESD's with over 65%, our local accuracy of the gridded surface, of their AIM plots mapping to a single ESG were recovered. These ESD's , `r unique(cleanNET4$ECO.SITE)`, were associated with `r nrow(cleanNET4)` AIM plots, and mapped to `r unique(cleanNET4$ESG)`. At this point the classified ESD's and non-classified ESD's appeared as Figures 3 and 4 respectively. 

```{r Plot Network of cleaning data, fig.cap="Relationships Between ESDs and ESGs midway through the cleaning process"}
par(mfrow=c(2,1)) 
plotweb(clean_mod@moduleWeb, labsize = 0.75,  col.high = '#1C315E', col.low = '#227C70', col.interaction = '#E6E2C3')
plotweb(dirt_mod@moduleWeb, labsize = 0.75,  col.high = '#1C315E', col.low = '#227C70', col.interaction = '#E6E2C3')
par(mfrow=c(1,1))
```

The remaining `r nrow(sgu)` ESD's, associated with `r nrow(ecosites2step2)` AIM Plots,  were all mapped to `r length(unique(sgu$ESG))` ESGs via the methods of summarizing both all three SGU layers in conjunction with climate zones. The final lookup table of ESD to ESG mapping is in Figure 4.  

```{r Plot Network of cleaned data a lookup table, fig.cap="Relationship between ESDs and ESGs at end of process"}
plotweb(final_mod@moduleWeb, labsize = 0.75, col.high = '#1C315E', col.low = '#227C70', col.interaction = '#E6E2C3')
rm(orig_mod, clean_mod, dirt_mod, final_mod)
```

```{r}
rm(known, prop_D_map2_G, number_correct_esg, number_incorrect, p2p, cleanNET1, 
   cleanNET2, cleanNET3, cleanNET4, sgu, ESG_present, ref_lookup)
```


# Conclusions

The relationship between the novel ESG's and the ESD's...

For best accuracy using the ESG's it is best practice to verify the ESD using traditional methods rather than spatial operations. 


\newpage

# References

